---
title: "CellID Wrapper Test 3"
author: "Nicolás Méndez"
date: "7/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# devtools::document()
# devtools::load_all(recompile = T)
# devtools::install()
# devtools::check()
# devtools::build()

library(tidyverse)
library(rcell2)

path <- "../data/image_samples/"
path.pdata <- paste0(path, "pdata.csv")
parameters <-  paste0(path, "parameters.txt")


difference <- function(lista) abs(lista[1] - lista[2])
```


https://web.archive.org/web/20190831012258/http://www.libtiff.org/index.html

http://gnuwin32.sourceforge.net/packages/tiff.htm

https://stackoverflow.com/a/57677953

https://stackoverflow.com/a/1618618

http://mingw-w64.org/doku.php/download

https://cygwin.com/install.html elegir todo lo que diga tiff con fuente e instalar cygwin


https://medium.com/@meghamohan/all-about-static-libraries-in-c-cea57990c495

https://www.geeksforgeeks.org/static-vs-dynamic-libraries/



https://stackoverflow.com/questions/58156585/build-rocksdb-static-library-inside-r-package

https://stackoverflow.com/questions/53631025/rcpp-install-package-with-static-libraries-for-platform-independent-usage

https://github.com/rwinlib/libtiff

https://stackoverflow.com/questions/53631025/rcpp-install-package-with-static-libraries-for-platform-independent-usage

https://stackoverflow.com/questions/58808393/how-do-i-create-an-r-package-that-depends-on-a-shared-library-with-rcpp

"As will the simpler approach of maybe just putting the source files of libbcd into the package src/ directory -- and R will take care of the rest."

Makevars.win

```
CC=ccache clang -Qunused-arguments
CXX=ccache clang++ -Qunused-arguments
CCACHE_CPP2=yes

PKG_CPPFLAGS =  -I"C:\Program Files (x86)\GnuWin32\include"
PKG_LIBS = -L"C:\Program Files (x86)\GnuWin32\lib" -ltiff
```

```
CC=gcc-4.4
CXX=g++-4.4

PKG_CXXFLAGS = -fno-keep-inline-dllexport
PKG_CFLAGS = -I"C:\cygwin64\usr\x86_64-w64-mingw32\sys-root\mingw\include" 
PKG_CPPFLAGS = -I"C:\cygwin64\usr\x86_64-w64-mingw32\sys-root\mingw\include"
PKG_LIBS = -L"C:\cygwin64\usr\x86_64-w64-mingw32\sys-root\mingw\lib" -ltiff
```

```
CC=ccache clang -Qunused-arguments
CXX=ccache clang++ -Qunused-arguments
CCACHE_CPP2=yes

PKG_CPPFLAGS =  -I"C:\Program Files (x86)\GnuWin32\include"
PKG_LIBS = -L"C:\Users\naikymen\Documents\R\win-library\4.0\tiff\libs\x64" -ltiff
```

# USAGE 0 - Exectuion arguments

```{r}
# Generate CellID input paths
cell.args <- cellArgs(#path = path,
                      path = "/run/media/nicomic/ACLN1/ACL/Uscope/datos_RtCC/2019-Dai-Far1-NG_reentry/Far1-NG/2019.12.12_Timecourse_FAR1-NG_RtCC/split",
                      parameters = parameters)
cellArgs.print(cell.args)  # Para revisar inputs
```

## Time series arguments

En cada archivo tiene que haber una lista con todos los tiempos para una sola posición.

```
cell -b /home/nicomic/split/Position01/bf_vcellid.txt -f /home/nicomic/split/Position01/fl_vcellid.txt -o /home/nicomic/split/Position01/out -p /home/nicomic/split/Position01/parameters_vcellid_out.txt
```

En `bf_vcellid.txt` solo van los BF:

    /home/nicomic/split/BF_Position01_time01.tif
    /home/nicomic/split/BF_Position01_time02.tif
    /home/nicomic/split/BF_Position01_time03.tif

En `fl_vcellid.txt` se meten todos los canalas de fluorescencia:

    /home/nicomic/split/YFP_Position01_time01.tif
    /home/nicomic/split/YFP_Position01_time02.tif
    /home/nicomic/split/YFP_Position01_time03.tif

El argumento de "out" y de los parámetos no cambia:

`parameters_vcellid_out.txt`

     max_split_over_minor 0.66
     max_dist_over_waist 8.00
     max_pixels_per_cell 1500
     min_pixels_per_cell 75
     background_reject_factor 1.00
     tracking_comparison 0.20
     image_type brightfield
     bf_fl_mapping list
    


```{r}
#
```


# USAGE 1 - RUN CELL ID


```{r}
# cell(cell.args, path = path, cell.command = "~/Software/cellID-linux/cell", )
# debugonce(cell)
cell(cell.args = cell.args,
     path = path,
     # cell.command = "~/Software/cellID-linux/cell",
     cell.command = "cellBUILTIN",
     no_cores = 3)
```


```{r}
cell.data <- cell.load(path = path,
                       pdata = path.pdata)

glimpse(cell.data)

# git checkout b8a1d7c3a760a4b373422d4ff1905cc193489901 .
# saveRDS(cell.data, "wapper_test.vcell.RDS")     # vcellid tests

# git checkout a6312b2ba1a4c53ea01ab4818d46099deb766073 .
# saveRDS(cell.data, "wapper_test.internal.RDS")  # internal cellid wrapper

# git checkout 0f44e6f7fa61c0158c9cd86b2dd738c0f39f60ce .
# saveRDS(cell.data, "wapper_test.external.RDS")  # external cellid wrapper
```

## Testing differences

All have used the same parameters.

```{r}
test <- cell.data$d

vcellstd <- readRDS("wapper_test.vcell.RDS")$d     # vcellid tests
internal <- readRDS("wapper_test.internal.RDS")$d  # internal cellid wrapper
external <- readRDS("wapper_test.external.RDS")$d  # external cellid wrapper
```


```{r}
nrow(test)
nrow(vcellstd)
nrow(internal)
nrow(external)
```


```{r}
dplyr::all_equal(test, vcellstd)

dplyr::all_equal(internal, external)
dplyr::all_equal(vcellstd, external)
dplyr::all_equal(internal, vcellstd)
```

```{r}
res <- compare::compareIgnoreOrder(vcellstd[,c("pos","xpos","ypos")], external[,c("pos","xpos","ypos")])

setdiff(vcellstd[,c("pos","xpos","ypos")], external[,c("pos","xpos","ypos")])


bind_rows(vcellstd, external, .id = "origin") %>% arrange(pos, xpos, ypos) %>% View()

d <- bind_rows(vcellstd, test, .id = "origin")

d$.duplicated <- bind_rows(vcellstd, test) %>% duplicated()

# d %>% filter(!.duplicated) %>% View()

difference <- function(a) abs(a[1] - a[2])

d %>% group_by(pos, xpos, ypos) %>% summarise(count = n(), 
                                              origin = paste(origin, collapse = ","),
                                              ucids =  paste(ucid, collapse = ","),
                                              ucids_diff = difference(as.numeric(ucid))) %>% View()

d %>% group_by(pos, xpos, ypos) %>% summarise(count = n(), 
                                              origin = paste(origin, collapse = ","),
                                              ucids =  paste(ucid, collapse = ","),
                                              ucids_diff = difference(as.numeric(ucid))) %>%
  filter(count != 2)
```


# USAGE 2

```{r}
#### SETUP PATHS DATAFRAME ####
paths <- magickPaths(cell.data)

paths
```

## USAGE 2.1 - Plot something

```{r}
cell.data$d %>%
  ggplot() +
  geom_point(aes(x = a.tot, y = fft.stat, color = pos)) +
  facet_wrap(~pos) +
  theme(legend.position = "none")

cell.data$d %>%
  ggplot() +
  geom_density(aes(x = f.tot.t, color = pos)) +
  theme_minimal()
```

## USAGE 2.2 - Pictures and filters

```{r}
cdata <- cell.data$d %>% 
  filter(ucid %in% c(
    2000000029
  ))
pdata <- read_csv(path.pdata)

magickCell(cdata, paths) %>% 
  magickForKnitr() %>% 
  knitr::include_graphics()
```

Slightly more useful:

```{r}
saved_data <- shinyCell(cell.data$d, pdata, paths,
                        plotType = "Dots",
                        # initial_facet = "pos ~ .",
                        #filters = saved_data$filters,
                        facet_grid_option = TRUE,
                        n_max = 7^2, max_size = 720)

# More plots would be useful
ggplot(data = subset(saved_data$cdata, filter == T)) +
  geom_boxplot(aes(y = cf.y, x = factor(pos)), width=0.1) +
  geom_violin(aes(y = cf.y, x = factor(pos), color = factor(pos)), alpha = 0) +
  theme_minimal()
```


```{r}
saved_data <- shinyCell(cell.data$d, pdata, paths,
                        plotType = "Dots",
                        # initial_facet = "pos ~ .",
                        filters = saved_data$filters,
                        facet_grid_option = TRUE,
                        n_max = 7^2, max_size = 720)

ggplot(data = subset(saved_data$cdata, filter == T)) +
  geom_point(aes(x=a.tot, y=el.p, color=factor(pos))) +
  theme_minimal()
```

## USAGE 2.3 - Pictures and filters reprise

```{r}
# Otra prueba
saved_data2 <- shinyCell(subset(saved_data$cdata, filter == TRUE), pdata, paths,
                         plotType = "Dots",
                         #initial_facet = "pH ~ treatment",
                         filters = saved_data$filters,
                         facet_grid_option = T,
                         n = 49, max_size = 720)

ggplot(data = subset(saved_data2$cdata, filter == T)) +
  geom_density(aes(x = a.tot, y=..scaled.., group = target), width=0.1) +
  # geom_violin(aes(y = f.tot.r, x = factor(target)), alpha = 0) +
  theme_minimal() + facet_wrap(~target)

ggplot(data = subset(saved_data2$cdata, filter == T)) +
  geom_boxplot(aes(y = f.tot.r, x = factor(target)), width=0.1) +
  # geom_violin(aes(y = f.tot.r, x = factor(target)), alpha = 0) +
  theme_minimal()

ggplot(data = subset(saved_data2$cdata, filter == T)) +
  geom_boxplot(aes(y = f.tot.y, x = factor(target)), width=0.1) +
  # geom_violin(aes(y = f.tot.r, x = factor(target)), alpha = 0) +
  theme_minimal()
```

