---
title: "CellID Wrapper Test 3"
author: "Nicolás Méndez"
date: "7/20/2020"
output: html_document
---

# Rebuild

```{r build, eval=FALSE, message=FALSE, warning=TRUE}
devtools::document()
devtools::load_all(recompile = T, reset = T)
devtools::install(reload = T, build = T,
                  quiet = T,
                  upgrade = "never", keep_source = T)
```

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rcell2)

difference <- function(lista) abs(lista[1] - lista[2])
```

# Run cellid mask_mod

## Con series temporales

```{r}
path <- "../data/time_seres_images_hu_moments/"
parameters <- "../inst/parameters.txt"
```

## Generate CellID input paths

```{r}
cell.args <- rcell2::cellArgs(path = path,
                              parameters = parameters,
                              BF.pattern = "BF_Position\\d+_time\\d+\\.tif$",
                              FP.pattern = "FP_Position\\d*_time\\d+\\.tif$",
                              O.pattern = ".*(Position\\d+)_time\\d+\\.tif"
                              )

rcell2::cellArgs.print(cell.args)  # Para revisar inputs
```

## Run cellid of the mask_mod branch

```{r, eval = F}
rcell2::cell(cell.args = cell.args,
             cell.command = "../inst/cellid_mask_mod",
             position.time.pattern = "time\\d+",
             no_cores = 1, dry = F)
```

# Load cell data

```{r}
path <- "../data/time_seres_images_hu_moments/"

# cell.data.old <- rcell2::cell.load.alt(path =  path)
cell_data <- rcell2::load_cell_data(path = path)
```

```{r}
cdata <- cell_data$data
paths <- cell_data$images %>% 
    mutate(file = paste0(path, "/", image))
```


# Load images

From `pos 1` and `t.frame 0`:

```{r}
pic_path <- paths %>% filter(pos == 1, t.frame == 0, channel == "BF.out")

pic_path$image
```

Also filter cdata for posterior comparison:

```{r}
cdata_filtered <- filter(cdata, pos == 1, t.frame == 0)
```

Read the image and plot it:

```{r}
pic <- tiff::readTIFF(pic_path$file)
pic_numeric <- pic * (2^16 - 1)

heatmap(pic_numeric[50:150,50:150], Rowv = NA, Colv = NA, scale = "none")
```
# Replace background with NA

```{r}
pic_numeric0 <- pic_numeric
pic_numeric0[pic_numeric0 == 0] <- NA

heatmap(pic_numeric0[50:150,50:150], Rowv = NA, Colv = NA, scale = "none")
```
# Convert to dataframe

## as.data.frame

```{r}
pic_df <- as.data.frame(pic_numeric0)

names(pic_df) <- 1:ncol(pic_df)

head(pic_df)
```
## reshape2::melt and manual method

```{r}
d1 <- reshape2::melt(pic_numeric) %>% 
  dplyr::rename(y = Var1, x = Var2)
# as.table(pic_numeric0)
d2 <- data.frame(y=rep(1:nrow(pic_numeric), ncol(pic_numeric)), 
                 x=rep(1:ncol(pic_numeric), each = nrow(pic_numeric)),
                 value=c(pic_numeric))

d1
d2

if(identical(d1, d2)) "Dataframes are identical!"
```
## Use manual method

```{r}
pic_df <- data.frame(y=rep(1:nrow(pic_numeric0), ncol(pic_numeric0)), 
                     x=rep(1:ncol(pic_numeric0), each = nrow(pic_numeric0)),
                     value=c(pic_numeric0))

pic_df <- pic_df[!is.na(pic_df[["value"]]),]

pic_df
```

# Summarise mask data

```{r}
pic_df_summary <- pic_df %>% 
  group_by(value) %>% 
  summarise(n_values = n(),
            # xspan = paste(min(x), max(x), sep = "-"),
            # yspan = paste(min(y), max(y), sep = "-"),
            xspan = min(x) - max(x),
            yspan = min(y) - max(y),
            mean_xpos = round(mean(x)),
            mean_ypos = round(mean(y))
            )

pic_df_summary
```


# Convert intensity to CellID

```{r}
bits <- 16

pic_df <- pic_df %>% 
  mutate(cellID = (2^bits -1) - value) %>% 
  arrange(cellID)

pic_df$cellID %>% unique()
```

```{r}
bits <- 16

pic_cdata <- pic_df_summary %>% 
  mutate(cellID = (2^bits -1) - value) %>% 
  arrange(cellID)

pic_cdata
```


# Compare with cdata

```{r}
cdata_compare <- cdata_filtered %>% 
  left_join(select(pic_cdata, 
                   cellID, mean_xpos, mean_ypos), 
            by = "cellID")

cdata_compare %>% 
  ggplot() +
    geom_point(aes(x = xpos, y = mean_xpos, color = factor(xpos - mean_xpos))) +
    xlab("cellid xpos") +
    ylab("rounded mean mask xpos")

cdata_compare %>% 
  ggplot() +
    geom_point(aes(x = ypos, y = mean_ypos, color = factor(ypos - mean_ypos))) +
    xlab("cellid ypos") +
    ylab("rounded mean mask ypos")
```

# Test Hu moments function

```{r}
one_cell.mask <- pic_df %>% nest(-cellID) %>% sample_n(1) %>% unnest()

one_cell.mask
```

```{r}
xy <- one_cell.mask %>% 
  select(x, y) %>% 
  as.matrix()

colnames(xy) <- c("dim1", "dim2")

xy %>% head()
```

```{r}
# source("../R/hu_moments_short.R")

hu.moments(xy)
```

## Run for all cells and append to cdata

# All-in-one function

```{r}
# source("../R/hu_moments_short.R")

pic_df_from_tiff <- function(tiff_path, image_bits){
    # Read masks tiff
    pic <- tiff::readTIFF(tiff_path)
    # Convert intensity value to 16-bit integers
    pic <- pic * (2^image_bits - 1)
    # Replace zeros with NA
    pic[pic == 0] <- NA
    
    # Convert pic matrix to dataframe of coordinates per mask/value
    pic_df <- data.frame(y=rep(1:nrow(pic), ncol(pic)), 
                         x=rep(1:ncol(pic), each = nrow(pic)),
                         pix_value=c(pic))
    # Clear NA rows
    pic_df <- pic_df[!is.na(pic_df[["pix_value"]]),]
    # Convert integer intensity value to CellID
    pic_df$cellID <- (2^image_bits -1) - pic_df[["pix_value"]]
    
    return(pic_df)
}

hues_from_xy <-  function(pic_df){
  # Split the dataframe by cellID
  pic_df_split <- pic_df %>% mutate(cellID = as.factor(cellID))
  pic_df_split <- split(pic_df_split, pic_df_split$cellID)
  
  # Compute Hu moments for each cellID's boundary mask XY coordinates
  hues <- lapply(pic_df_split, FUN = function(cell_coords_df){
    # Convert dataframe to matrix
    xy <- cell_coords_df %>% 
          select(x, y) %>% 
          as.matrix()
    # Rename XY columns appropriately
    colnames(xy) <- c("dim1", "dim2")
    
    # Return a named vector with the cell ids and the named hu moments
    c(cellID = unique(cell_coords_df[["cellID"]]), 
      hu.moments(xy))
  })
  
  # Bind rows from all cellIDs and return
  return(bind_rows(hues))
}

append_hues <- function(cell_data, image_bits = 16, return_points = F){
  paths <- cell_data$images %>% 
    mutate(file = paste0(path, "/", image)) %>% 
    filter(channel == "BF.out")
  
  # For testing
  # pic_metadata <- c()
  # pic_metadata["file"] <- paths[1,]$file
  # pic_metadata["t.frame"] <- paths[1,]$t.frame
  # pic_metadata["pos"] <- paths[1,]$pos
  # image_bits = 16
  # return_points = T
  
  pic.and.hues.dfs <- apply(paths, MARGIN = 1, FUN = function(pic_metadata){
    # Extract xy coordinates list from the BF.out mask tiff 
    pic_df <- pic_df_from_tiff(tiff_path = pic_metadata["file"], 
                               image_bits = image_bits)
    
    # Compute Hu moments for each cellID
    hues_df <- hues_from_xy(pic_df)
    
    # Add position and time information
    hues_df$t.frame <- as.integer(pic_metadata["t.frame"])
    hues_df$pos <- as.integer(pic_metadata["pos"])
    pic_df$t.frame <- as.integer(pic_metadata["t.frame"])
    pic_df$pos <- as.integer(pic_metadata["pos"])
    
    # Return the Hu moments and points dataframes
    list(pic_df = if(return_points) pic_df else NULL, 
         hues_df = hues_df)
  })
  
  # Bind hues dataframes from all pics
  hues.df <- bind_rows(
    lapply(pic.and.hues.dfs, FUN = function(l) l[["hues_df"]])
  )
  
  # Append Hu moment data to cell_data object
  cell_data[["data"]] <- left_join(cell_data[["data"]],
                                   hues.df, 
                                   by = c("cellID", "t.frame", "pos"))
  
  if(return_points){
    masks_df <- bind_rows(
      lapply(pic.and.hues.dfs, FUN = function(l) l[["pic_df"]])
    )
    
    cell_data[["masks"]] <- masks_df
  }
  
  return(cell_data)
}
```

## Run

```{r}
cell_data2 <- append_hues(cell_data, return_points = T)
```

## Check

```{r}
masks_df_summary <- cell_data2$masks %>% 
  group_by(cellID, t.frame, pos) %>% 
  summarise(cellID = first(cellID),
            t.frame = first(t.frame),
            pos = first(pos),
            n_values = n(),
            # xspan = paste(min(x), max(x), sep = "-"),
            # yspan = paste(min(y), max(y), sep = "-"),
            xspan = min(x) - max(x),
            yspan = min(y) - max(y),
            mean_xpos = round(mean(x)),
            mean_ypos = round(mean(y))
            )

masks_df_summary
```
```{r}
cell_data2$masks$cellID %>% range()
cell_data2$data$cellID %>% range()
```


```{r}
cdata_compare <- cell_data2$data %>% 
  left_join(select(masks_df_summary, 
                   cellID, t.frame, pos, mean_xpos, mean_ypos), 
            by = c("cellID", "t.frame", "pos"))

cdata_compare %>% 
  ggplot() +
    geom_point(aes(x = xpos, y = mean_xpos, color = factor(xpos - mean_xpos))) +
    facet_grid(pos~t.frame) +
    xlab("cellid xpos") +
    ylab("rounded mean mask xpos")

cdata_compare %>% 
  ggplot() +
    geom_point(aes(x = ypos, y = mean_ypos, color = factor(ypos - mean_ypos))) +
    facet_grid(pos~t.frame) +
    xlab("cellid ypos") +
    ylab("rounded mean mask ypos")
```

