---
title: "CellID Wrapper Test 3"
author: "Nicolás Méndez"
date: "7/20/2020"
output: html_document
---

# Rebuild

```{r build, message=FALSE, warning=TRUE}
devtools::document()
devtools::load_all(recompile = F, reset = T, quiet = T)
devtools::install(reload = T, build = F,
                  quiet = T,
                  upgrade = "never", keep_source = T)
# devtools::install_github("gerbeldo/rcell2", ref = "rcellid_no_labels")
# devtools::check()
# devtools::build()
# devtools::build(path="../../test_libtiff.tar.gz")
# remove.packages("rcell2")
# install.packages("../../test_libtiff.tar.gz", repos = NULL, type="source", clean = TRUE)
```

```{r setup, message=FALSE, warning=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
# library(rcell2)

path <- "../data/image_samples/"
path.pdata <- paste0(path, "pdata.csv")
parameters <- "../inst/parameters.txt"

difference <- function(lista) abs(lista[1] - lista[2])
```

## Reinstall

```{r, eval=F}
devtools::install(reload = T, build = T,
                  quiet = T,
                  upgrade = "never", keep_source = T)
```


# USAGE

## Load data and plot something

```{r}
cell.data <- rcell2::cell.load.alt(path =  path, ucid.zero.pad = 5)

head(cell.data$d) %>% select(ucid, pos, cellID, t.frame)

#### SETUP PATHS DATAFRAME ####
paths <- rcell2::magickPaths(cell.data)

p <- cell.data$d %>% ggplot() + 
  geom_point(aes(x = a.tot, y = fft.stat, color = pos)) +
  theme(legend.position = "none")

p
```

# Test shinyCell

```{r}
cdata <- cell.data$d
```

## Basic usage

```{r}
saved_data <- shinyCell(cdata, pdata, paths,
                        plotType = "Dots",
                        # initial_facet = "pos ~ .",
                        #filters = saved_data$filters,
                        facet_grid_option = TRUE,
                        n_max = 7^2, max_size = 720)

# More plots would be useful
ggplot(data = subset(saved_data$cdata, filter == T)) +
  geom_boxplot(aes(y = cf.y, x = factor(pos)), width=0.1) +
  geom_violin(aes(y = cf.y, x = factor(pos), color = factor(pos)), alpha = 0) +
  theme_minimal()
```

## Display and play with filters and polygon

You may combine several filter lists:

```{r}
poly.filters <- c(
  saved_data$filters
)
```

Time-series cdata must have a unique id column:

```{r}
cdata.ucidtime <- cdata %>% mutate(id = paste(ucid, t.frame, sep = "_"))
```

Then, any filter list can be used to filter any cdata-like dataframe:

```{r}
cdata.ucidtime.filters <- polyFilterApply(polygon_df_list = poly.filters, 
                                          cdata = cdata.ucidtime, 
                                          cell_unique_id_field = "id")
```

Filters can be plotted like this:

```{r}
plot_filters(saved_data)
```


### Bind polygon filter lists

```{r}
saved_data <- readRDS("data/20191212_Timecourse_FAR1-NG_RtCC/output/objetos_intermedios/shiny_saved_data_post_k1.RDS")
saved_data2 <- readRDS("data/20191212_Timecourse_FAR1-NG_RtCC/output/objetos_intermedios/shiny_saved_data2_post_k1.RDS")
```

```{r}
poly.filters.list <- c(
  saved_data$filters,
  saved_data2$filters
)
poly.filters.list.names <- 1:length(poly.filters.list)
```

```{r}
bfs <- bind_filters(list(cdata = saved_data$cdata, filters = poly.filters.list))

plot_bound_filters(bfs)
```

### Find polygon intersections

```{r}
devtools::load_all("/home/nicomic/Projects/Rdevel/rcell2/", reset = T)
```

Requires `rgeos` (for the intersection function) and `sp` (as an input dependency of the former):

```{r}
list.sp.poly <- lapply(poly.filters, function(poly) sp::Polygon(poly[,c("x", "y")], hole = F))

sp.poly.one <- sp::Polygons(list.sp.poly[1], ID = "first")
sp.poly.one <- sp::SpatialPolygons(list(sp.poly.one))

sp.poly.two <- sp::Polygons(list.sp.poly[2], ID = "second")
sp.poly.two <- sp::SpatialPolygons(list(sp.poly.two))

sp.poly.three <- sp::Polygons(list.sp.poly[3], ID = "third")
sp.poly.three <- sp::SpatialPolygons(list(sp.poly.three))
```


```{r}
plot_filters(list(cdata = saved_data$cdata, filters = poly.filters.list[c(1,2,3)]))
```

No intersection returns `NULL`:

```{r}
rgeos.intersect <- rgeos::gIntersection(spgeom1 = sp.poly.one, sp.poly.three)

rgeos.intersect
```

An intersection returns an `sp` object (or something like that):

```{r}
rgeos.intersect <- rgeos::gIntersection(spgeom1 = sp.poly.one, sp.poly.two)

rgeos.intersect
```

The intersection coords are here:

```{r}
intersect.coords <- rgeos.intersect@polygons[[1]]@Polygons[[1]]@coords
```

Notice the appearance of a fourth polygon:

```{r}
intersect.coords.df <- as.data.frame(intersect.coords)

intersect.coords.df$xvar <- "a.tot"
intersect.coords.df$yvar <- "el.p"
intersect.coords.df$type <- "Subtractive"

# ggplot(as.data.frame(intersect.coords)) + geom_polygon(aes(x=x,y=y))
plot_filters(list(cdata = saved_data$cdata, filters = c(poly.filters.list[c(1,2,3)],
                                                        list(intersect.coords.df))))
```
