---
title: "Rcell2 Workflow Template"
author: "Nicolás Méndez"
date: "30/09/2021"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: false
    toc_depth: 4
    number_sections: false
    smooth_scroll: false
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
urlcolor: blue
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "output/renders") })
bibliography: references.bib
---

<!-- Download a copy of this file with rcell2::get_workflow_template() -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magick)  # if you need it
library(utiles)  # if you need it, get this from: https://gitlab.com/naikymen/r_utils
library(rcell2)  # this package you are checking out :)
library(tidyverse)  # import numpy as np (?)
```

## Friendly reminders

**READ THE README**.md at rcell2's GitHub repo.

**MIRAR LAS FOTOS** CON IMAGEJ ANTES DE EMPEZAR. Te puede ahorrar mucho tiempo :D

## FULL EXAMPLE (TL;DR)

Run cellid using the selected parameters.

Otherwise you may omit `parameters = parameters.txt` in the `rcell2::arguments()` call,
or set your own using `rcell2::parameters.default()` and `rcell2::parameters.write()`.

```{r}
data.dir <- "data/image_samples/"

parameters.txt <- rcell2::parameters.write(rcell2::parameters.default())

cellid.args <- rcell2::arguments(path = data.dir,
                                 parameters = parameters.txt,  # optional
                                 file.pattern = "^(BF|TFP)_Position(\\d+)().tif$")
cell2.out <- 
  rcell2::cell2(arguments = cellid.args, 
                output_coords_to_tsv = T,  # Optional
                cell.command = "~/Software/cellID-linux/cell")

cell.data <- rcell2::cell.load.alt(path = data.dir,
                                   fluorescence.pattern = "^([GCYRT]FP|[GCYRT]\\d+)_Position\\d+.tif$")

cell.boundaries <- 
  rcell2::cell.load.boundaries(data.source = "masks.tsv",
                               arguments = cellid.args,
                               pixel.type = "b"
                               )
```

### Guardar todo el output

Puede ser conveniente guardar todo en una lista:

```{r}
saveRDS(
  file = paste0(data.dir, "/", "cell2.objects.RDS"),
  object = list(
    parameters.df=parameters.df[2,],
    cellid.args=cellid.args,
    cell2.out=cell2.out,
    cell.data=cell.data,
    cell.boundaries=cell.boundaries
  ))
```

### Restaurar output

La listas se pueden cargar como objetos en el environment así: https://stackoverflow.com/a/13797968

```{r}
list2env(readRDS(paste0(data.dir, "/", "cell2.objects.RDS")),
         globalenv())
```

## Renaming Metamorph MDA images

Using Metamorph's _Multi-dimensional acquisition_ (MDA for short), may use file names different from the usual style (i.e `YFP_Position01_time1.tif`, and so on).

`rename.mda` will help you renaming those weird file names into something more useful.

```{r}
images.path <- "~/Projects/PhD/data/uscope/multidimensional_exp-20211126-Far1NG-wt_y_dKar4/"
rename.mda(images.path)
```

See `?rename.mda` for extra options.

The defaults are useful for a three channel experiment, and file names looking like this: `far1_rtcc_exp16_thumb_w1LED-BF--YFPcube--cam_s17_t35.TIF`.

```{r}
formals(rename.mda)
```

These are easily adaptable to other use cases, for example you may change `channel.dict` to include other channels, in a different order.

Note that the `ch` field must match the corresponding capture group in `identifier.pattern` exactly.

## CellID Examples

### Parameters for CellID

Description for parameters:

```{r}
rcell2::cellid_parameter_descriptions(list_format = F)
```

A prototype parameter list, with defaults:

```{r}
parameters.list <- list(
  max_split_over_minor = 0.50,
  max_dist_over_waist = 8.00,
  max_pixels_per_cell = 2000,
  min_pixels_per_cell = 75,
  background_reject_factor = 0.75,
  tracking_comparison = 0.20,
  align_individual_cells = F,
  align_fl_to_bf = T,
  image_type = "brightfield",
  bf_fl_mapping = "list"
  )
```

There is a helper function to make one more easily:

```{r}
parameters.list <- rcell2::parameters.default()

parameters.list$max_split_over_minor
```

Hay que guardar los parámetros en un archivo `parameters.txt`,
para que más adelante CellID lo pueda usar.

Para eso está la siguiente función, 
que devuelve el path donde se guardó el archivo:

```{r}
parameters.txt <- rcell2::parameters.write(parameters.list)

parameters.txt
```

Revisemos el contenido del `parameters.txt` que generamos recién:

```{r}
writeLines(readLines(parameters.txt))
```

#### Test many!

A veces querés probar muchos parámetros antes de correr CellID en todas las fotos.

For each parameter variation, we need a new parameters file.

In this example, only `background_reject_factor` is changed:

```{r}
parameters.list <- rcell2::parameters.default(
  background_reject_factor = seq(from=0, to=1.5, by = 0.6)
)

parameters.list$background_reject_factor
```

If tou have my `utiles` package, this may be useful too:

```{r}
parameters.list <- rcell2::parameters.default()

parameters.list$background_reject_factor <-
  utiles::value.to.range(
    parameters.list$background_reject_factor,
    width = 0.75, by = 0.1)

parameters.list$background_reject_factor
```

The list must now be expanded into a data.frame:

```{r}
parameters.df <- expand.grid(parameters.list)

test.params <- 1:nrow(parameters.df)  # Record the amount of combinations

parameters.df
```

> Note: You might want to select some rows from that table, in that case, simply filter it.

> Note: `utiles::value.to.range` just converts a number to a vector centered on its value (see `base::seq`).

Define working paths for parameter testing:

* One with the images.
* Another one for testing.

```{r}
test.dir <- "/tmp/images_directory/test.dir"
unlink(test.dir, recursive = T)
dir.create(test.dir, recursive = T)
```

Define the path of the original images:

```{r}
data.dir <- "data/image_samples/"
```

Check that their file names are properly recognized:

```{r}
cellid.args <- rcell2::arguments(data.dir,
                                 file.pattern = "^(BF|TFP)_Position(\\d+)().tif$")

cellid.args
```

Now choose some *positions* (image sets) and times to use as test for each parameter set:

```{r}
test.pos = 1:2
test.frames = 1

subset(cellid.args,
       pos %in% test.pos & t.frame %in% test.frames)
```

Now comes the magic part. Ut uses temporary directories to work on, 
and avoids messing up the main data directory.

Remember to adjust file patterns in `rcell2::arguments` and `rcell2::cell.load.alt`.

Original images file pattern:

```{r}
# Default value: all channels, no z-stack.
arguments.file.pattern <- "^(BF|[A-Z]FP)_Position(\\d+)_time(\\d+).tif$"

# BF or TFP only, no t.frame info.
arguments.file.pattern <- "^(BF|TFP)_Position(\\d+)().tif$"
```

Fluorescent channel pattern:

```{r}
# Default value: all fluorescent tags, including Z-stack fluorescence channels.
load.cell.channel.pattern <- "^([GCYRT]FP|[GCYRT]\\d+)_Position\\d+_time\\d+.tif$"
# Same as defualt, but without t.frame information.
load.cell.channel.pattern <- "^([GCYRT]FP|[GCYRT]\\d+)_Position\\d+.tif$"
```

And the CellID path:

```{r}
cell.command.path = "~/Software/cellID-linux/cell"
```

Create a parallel backend:

```{r}
ncores <- parallel::detectCores() - 1
cl <- parallel::makeCluster(ncores, setup_strategy = "sequential")
# doParallel::registerDoParallel(cl)
doSNOW::registerDoSNOW(cl)
```

Setup a progressbar and run CellID:

```{r}
ntasks <- length(test.params)
pb <- txtProgressBar(max = ntasks, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress=progress)

library(foreach)

result <- 
  foreach(test.param=test.params, #.export = c(".parameters.df",)
          .options.snow=opts,
          .packages = c("rcell2", "base", "dplyr")) %dopar% {
    
            # Get one parameter set
            parameters.list.one <- parameters.df[test.param,]
            
            # Save parameters
            parameters.txt <- rcell2::parameters.write(parameters.list.one)
            
            # Create arguments
            cellid.args <- rcell2::arguments(data.dir,
                                             file.pattern = arguments.file.pattern,
                                             parameters = parameters.txt)
            # Subset only one position
            # cellid.args.one <- cellid.args[cellid.args$pos %in% test.pos, ]
            cellid.args.one <- subset(cellid.args, 
                                      pos %in% test.pos & t.frame %in% test.frames)
            
            # Prepare a temp dir for each cellid run
            tmp.path <- tempfile(pattern = "dir", tmpdir = test.dir)
            dir.create(tmp.path)
            
            # Create shortcuts to the original images
            apply(cellid.args.one, 1, function(i){
              file.symlink(from = paste0(i["path"], "/",
                                         c(i["image"],i["bf"])),
                           to = paste0(tmp.path, "/", 
                                       c(i["image"], i["bf"]))
                     )
            })
        
            
            # Regenerate arguments for the new tmp path
            cellid.args.tmp <- rcell2::arguments(tmp.path,
                                                 file.pattern = arguments.file.pattern,
                                                 parameters = parameters.txt)
            # Run cellid
            rcell2::cell2(arguments = cellid.args.tmp, 
                          cell.command = cell.command.path)
            
            # Load output
            cell.data <- rcell2::cell.load.alt(
              tmp.path,
              fluorescence.pattern = load.cell.channel.pattern
              )
            
            # Keep only the image paths dataframe
            new.images <- 
              cell.data$images %>% filter(is.out)
            
            return(new.images)
  }

close(pb)
parallel::stopCluster(cl)
```

Bind the result:

```{r}
results.bound <- result %>% 
    bind_rows(.id = "id") %>%
    mutate(id = as.integer(id))

results.bound
# saveRDS(results.bound, tempfile(pattern = "results", tmpdir = test.dir, fileext = ".RDS"))
```

Make TIFF stacks for each channel and cleanup:

```{r}
stack.paths <- results.bound %>% 
    arrange(channel, id, t.frame, pos) %>% split(~channel+pos) %>% 
    lapply(function(images){
      # images <- stack.paths[[1]]
      stack.name <- paste0(
          test.dir, "/", images$channel[1], "_stack-pos_", images$pos[1], #"-time", images$t.frame[1],
          ".tif"
          )
      
      images %>% 
        dplyr::arrange(channel, t.frame, pos, id) %>% 
        with(file) %>% 
        magick::image_read() %>% 
        magick::image_write(path=stack.name)
      
      return(stack.name)
    })

stack.paths  # The stacks are saved to these paths
```

Review output in imagej and choose a parameter set:

```{r}
parameters.list <- parameters.df[2,]
```

> TODO: nombrar los stacks por parameter modificado

BONUS: Leer fotitos con magick!

```{r}
bf.out.imgs.paths <- dir(path, pattern = "^BF.*.out.tif$", full.names = T)

bf.out.imgs <- magick::image_read(bf.out.imgs.paths) %>% 
  magick::image_annotate(round(bgrf.new,2), color = "white", size = 10, boxcolor = "black")

magick::image_append(bf.out.imgs, stack = T)
```

### Arguments for CellID

Es un dataframe con información para corerr CellID.

La opción más importante es `file.pattern` que es una expresión regular para encontrar archivos de imágenes en el `path`.

Por defecto es:

    "^(BF|[A-Z]FP)_Position(\\d+)_time(\\d+).tif$"

Entre paréntesis están los "grupos de captura" de la expresión regular, donde se espera que esté la siguiente información:

1. Canal
2. Número de posición
3. Número de tiempo

Para no usar tiempo solo hay que borrar la parte del tiempo, dejando un paréntesis vacío en su lugar `()`:

    "^(BF|[A-Z]FP)_Position(\\d+)().tif$"

Se puede "filtrar" posiciones y tiempos directo desde la regex, o posteriormente usando las funciones de R usuales. Para usar solo la posicion 2, y los tiempos 4, 2 y 3, la regex sería:

    "^(BF|[A-Z]FP)_Position(?0+2)_time(?0+[423]).tif$"

Noten el `0+?` enfrente del número, que indica que puede haber un cero o más frente al numerito de posición o tiempo en el nombre del archivo.

```{r}
path <- "~/Software/cellID-linux/samples/"

cellid.args <- rcell2::arguments(path = path, 
                                 file.pattern = "^(BF|[A-Z]FP)_Position(\\d+)_time(\\d+).tif$",
                                 parameters = parameters)
cellid.args
```

Obs: su `file.pattern` regex puede variar, por ejemplo si no hay tiempo y solo hay BFs y TFPs, usarian:

* `file.pattern = "^(BF|TFP)_Position(\\d+)().tif$"`

Nota: pueden ver que pasa filtrando con `dplyr`:

```{r}
cellid.args %>% filter(pos == 2, t.frame %in% c(4,2,3))
```


### Run CellID

La función `cell2` genera comandos para llamar a CellID por línea de comandos y los ejecuta en paralelo.

La opción `dry = T` indica a la función hacer "todo" menos ejecutar CellID, e imprime los comandos que habría usado.

Para correr los comandos, usen `dry = F` (aunque tambien pueden copiarlos y pegarlos en una terminal, si están en Linux o Mac OS).

El parámetro `cell.command` es el path al ejecutable. En ciertas versiones de `rcell2`,
el binario se compila durante la instalación y se usa por defecto.

```{r}
rcell2::cell2(arguments = cellid.args, 
              cell.command = "~/Software/cellID-linux/cell",
              dry = T, no_cores = 2)
```

## Load CellID data

Esto se hace con la función `load_cell_data`:

```{r}
cell.data <- rcell2::load_cell_data(path = path)
```

Si esa no funciona, pueden intentar con `cell.load.alt`:

```{r}
cell.data <- rcell2::cell.load.alt(path = path,
                                   fluorescence.pattern = "^([GCYRT]FP)_Position\\d+.tif$")
```

Obs: su `fluorescence.pattern` regex puede variar.

### CellID variables help

Descriptions of the variables in `cdata`:

```{r}
rcell2::cellid_output_descriptions()
```

### Check XY stage coordinates against expectations

It may be a good idea to double check that the physical coordinates of your images
match your expectations (i.e. the column/row pairs of each well in a microscopy plate).

This is specially useful when tracing a path for positions in a many-well time-course experiment.

The following code match the images' _position_ (Pos) to their microscope stage _physical_ XY coordinates,
which are stored in the TIFF tags of the original images.

It requires the `images` dataframe in `cell.data` to get the file names,
_but_ you could also use `dir` to get the files by pattern, and `sub` to get the position IDs,
independently of CellID.

```{r}
tags <- images %>% filter(t.frame==0, channel=="BF") %>% 
  {split(.$file, .$pos)} %>% 
  lapply(rcel2::read_tiff_tags)

# any(is.null(names(tags$`1`$frame1)))  # Chekc if any names are missing
# select.list(tags$`1`$frame1, discard = "description")
# tags$`1`$frame1$description

tags <- tags %>% 
  lapply(function(x) x[["frame1"]][["description"]]) %>% 
  bind_rows(.id = "pos") %>% 
  mutate(variable = make.names(variable)) %>% 
  filter(variable %in% make.names(c("stage-position-x", "stage-position-y", "z-position")))

tags <- tags %>%
  pivot_wider(id_cols = "pos", names_from = "variable", values_from = "value") %>% 
  mutate_all(as.numeric)
```

Then make a scatterplot:

```{r}
tags %>% arrange(pos) %>% 
  ggplot(aes(x=stage.position.x, y=stage.position.y, label=pos)) +
  geom_path() + 
  geom_text(size=10) +
  geom_point(color="red") +
  scale_x_reverse() +
  ggtitle("Coordenadas del stage")
```

## Analizar datos

### Tidy framework

- [ ] TODO: ¿donde puso los ejemplos/viñetas Ger para su tidy framework?

### Shiny-Magick framework

Inputs importantes:

```{r}
cdata <- cell.data$data
images <- cell.data$images  # puede tener otro nombre el data.frame con info de las fotos
pdata <- read.csv("data/pdata.csv")
```

Uso mínimo:

```{r}
shiny.filter <- shinyCell(cdata, pdata, images)
```

Si hay muchas celulas, conviene iniciar con `plotType = "Hex"` u otro plot aglomerativo.

Para ver las regiones filtradas, hay una función que grafica los filtros:

```{r}
plot_filters(shiny.filter, print_plots = F)
```

Ver ejemplos (un poco viejos) en:

* `testings/examples_celll_and_shiny.Rmd`
* `testings/test_magickCell.Rmd`
* `testings/test_shinyCell.Rmd`
* `testings/test_tagCell.Rmd`

- [ ] TODO: documentar esto bien.

### Supervised classification

- [ ] TODO: agregar ejemplos de funciones de los clasificadores entrenados (LR y SVM).

* `testings/linear_classification`
* `testings/tree_classification`

### Unsupervised classification

- [ ] TODO: agregar ejemplos de funciones de k-means.

* `testings/kmeans_filtering_ac`

## Datos de máscaras

Hay dos formas de cargar los datos de las mascaras, segun las opciones con las que corriste `cell2`.

### TIFF pixel ID

Si especificaron la opción `encode_cellID_in_pixels` en `cell2`, el CellID está codificado en la intensidad de pixel de las fotos de `BF.out`, y se puede cargar usando la funcion `pic_df_from_tiff`.

```{r}
mask <- 
  rcell2:::pic_df_from_tiff(tiff_path = "../data/image_samples/BF_Position001.tif.out.tif", 
                            image_bits = 16)
```

### TSV output

Si especificaron la opción `output_coords_to_tsv` en `cell2`, hay un archivo TSV por posición con las coordenadas de los bordes y del interior todas las células en esa posición.

Pueden cargarlo y revisarlo así:

```{r}
masks <- read_tsv("~/Software/cellID-linux/test_data_tmp/Position01/out_all_masks.tsv")

masks
```

Por ejemplo:

```{r}
p1 <- ggplot(masks) + geom_tile(aes(x=x, y=y)) + facet_grid(flag~pixtype) + ggtitle("Todas las celulas")
p2 <- masks %>% filter(cellID == 0, t.frame == 0) %>% ggplot() + geom_tile(aes(x=x, y=y)) + facet_grid(flag~pixtype) + ggtitle("Una celula")

p1
p2
```

#### Carga automatica

Esta función requiere el objeto `cellid.args` generado antes:

```{r}
cell.boundaries <- 
  rcell2::cell.load.boundaries(data.source = "masks.tsv",
                               arguments = cellid.args,
                               pixel.type = "b"
                               )
```


### Analisis de máscaras

#### Hu Moments

Usar datos de máscaras para calcular sus Hu Moments.

Para usar los datos de los TIFF, solo hace falta correr lo siguiente:

```{r}
cell_data <- rcell2::append_hues(cell_data = cell_data, 
                                 return_points = T, 
                                 image_bits = 16)

rcell2:::check_tiff_mask(cell_data)
```
 
Para usar los TFP, ver `testings/test_hu_moments.Rmd`.

## Other tools (WIP)

There are various mini-projects for analizyng boundary/mask/cellid data. See:

* `testings/edge_kernels`
* `testings/erosion_dilation`
* `testings/focus_scores`
* `testings/hu_moments`
* `testings/ijtiff_masks`
* `testings/kmeans_filtering_ac`
* `testings/linear_classification`
* `testings/noise_tests`
* `testings/tree_classification`
* `testings/yeast_vision`

### ImageJ FFT filter on BF images

* `testings/fft_filter`

### Cell shapes / Boundary angles

* `testings/curvature_tests`

- [ ] TODO: agregar ejemplos de funciones de suavizado y parametrizacion.
- [ ] TODO: agregar ejemplos de funciones de DTW para alineamiento.

### Signal localization

- [ ] TODO: agregar las cosas de SNR del curso de Luciana.
