---
title: "Edge erosion for Cell-ID masks"
---

```{r setup}
devtools::load_all("~/Projects/Academia/Colman/gitlabs_acl/utiles", reset = T)
# devtools::load_all("~/Projects/Academia/Colman/gitlabs_acl/rcell2", reset = T)
# devtools::install()
library(rcell2)
library(tidyverse)

knitr::opts_knit$set(root.dir = here::here())
```

> Ni idea de lo que estoy haciendo

## Datos

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix.tif", binarize = T)
mask.paded <- mat_pad(mask)

plot_binary_matrix(mask)
```


```{r}
# m33 <- matrix(c(1,0,0,1,0,0,0,0,0), 3)

plot_binary_matrix(mask[1:3, 5:7])
plot_binary_matrix(mask[5:7, 4:6])
plot_binary_matrix(mask[4:6, 2:4])
```

## Automata falopa

### Regla extremo doblado

Kill

```{r}
m33 <- matrix(c(0,0,0,0,1,0,1,1,0), 3)
plot_binary_matrix(m33)

check_bent_tip <- function(m33){
    
    # Ensure exactly 2 neighbors
    if(!(sum(m33) - 1) == 2) return(FALSE)
  
    # turn around the center, counter-clockwise (CCW), starting at [1,1]
    loop_sum <- c(1, 2, 3, 6, 9:7, 4)
    # m33[loop_3x3]
    
    # same, but lagged
    loop_3x3_lagged <- c(4, 1, 2, 3, 6, 9:7)
    # m33[loop_3x3_lagged]
    
    # Sum each element with the previous one:
    loop_sum <- m33[loop_sum] + m33[loop_3x3_lagged]
    
    # Loop should match exactly 
    pattern <- c(rep(0,5),1,2,1)
    loop_sum_cat <- c(loop_sum, loop_sum)
    
    # Find the pattern
    idxs.length <- length(loop_sum_cat) - length(pattern) + 1
    for(i in 1:idxs.length){
      
      loop_sum_cat.segment <- loop_sum_cat[i:(i-1+length(pattern))]
      
      if(all(loop_sum_cat.segment == pattern)){
        return(TRUE)
      }
    }
    
    return(FALSE)
}
```

### Regla protuberancia

Kill

```{r}
m33 <- matrix(c(1,1,1,0,1,0,0,0,0), 3)
plot_binary_matrix(m33)

check_protuberance <- function(m33){
  
  m33[2,2] <- 0
  
  col_condition <- all.equal(sort(colSums(m33)), c(0,0,3))
  
  row_condition <- all.equal(sort(rowSums(m33)), c(0,0,3))
  
  return(isTRUE(col_condition) | isTRUE(row_condition))
}

check_protuberance(m33)
```
### Square - esta bien?

Kill

```{r}
m33 <- matrix(c(0,0,0,1,1,0,1,1,0), 3)
# m33 <- matrix(c(1,1,1,0,1,0,0,0,0), 3)
plot_binary_matrix(m33)

check_square <- function(m33){
  
  resultado <- FALSE  

  # turn around the center, counter-clockwise (CCW), starting at [1,1]
  loop_3x3 <- c(1, 2, 3, 6, 9:7, 4)
  
  # Loop should match exactly 
  pattern <- c(0,0,0,0,0,1,1,1)
  loop_cat <- c(m33[loop_3x3], m33[loop_3x3])
  # loop_sum_cat
  
  # Find the pattern
  idxs.length <- length(loop_cat) - length(pattern) + 1
  for(i in 1:idxs.length){
    
    loop_cat.segment <- loop_cat[i:(i-1+length(pattern))]
    
    if(all(loop_cat.segment == pattern)){
      resultado <- TRUE
      break
    }
  }
  
  resultado <- resultado & any(colSums(m33) == 0) & any(rowSums(m33) == 0)
  
  return(resultado)
}

check_square(m33)
```

### Regla segmento

Keep

```{r}
m33 <- matrix(c(0,0,0,1,1,1,0,0,0), 3)
plot_binary_matrix(m33)

check_segment <- function(m33){
  
  resultado <- FALSE  

  # turn around the center, counter-clockwise (CCW), starting at [1,1]
  loop_3x3 <- c(1, 2, 3, 6, 9:7, 4)
  
  # Loop should match exactly 
  pattern <- c(1,0,0,0,1,0,0,0)
  loop_cat <- c(m33[loop_3x3], m33[loop_3x3])
  # loop_sum_cat
  
  # Find the pattern
  idxs.length <- length(loop_cat) - length(pattern) + 1
  for(i in 1:idxs.length){
    
    loop_cat.segment <- loop_cat[i:(i-1+length(pattern))]
    
    if(all(loop_cat.segment == pattern)){
      resultado <- TRUE
      break
    }
  }
  
  return(resultado)
}

check_segment(m33)
```

### Regla de la "L"

Keep

```{r}
m33 <- matrix(c(0,1,0,0,1,0,0,1,1), 3)
# m33 <- matrix(c(0,1,0,0,1,1,0,1,1), 3)
plot_binary_matrix(m33)

check_ele <- function(m33){
  
  m33[2,2] <- 0
  
  sum_condition <- all.equal(c(0,2) , sort(c(rowSums(m33)[2], colSums(m33)[2])))
  
  diag_condition <- m33[1,1] + m33[1,3] + m33[3,1] + m33[3,3] == 1
  
  return(isTRUE(sum_condition) & diag_condition)
}

check_ele(m33)
```

### Regla de la "S"

Keep

```{r}
m33 <- matrix(c(1,1,0,0,1,0,0,1,1), 3)
plot_binary_matrix(m33)

check_ese <- function(m33){
  
  m33[2,2] <- 0
  
  count_condition <- sum(m33) == 4
  
  sum_condition <- rowSums(m33)[2] == 2 | colSums(m33)[2] == 2
  
  diag_condition <- sum(m33[1,1], m33[3,3]) == 2 | sum(m33[3,1], m33[1,3]) == 2
  
  return(count_condition & sum_condition & diag_condition)
}

check_ese(m33)
```

### Regla del pescado

Keep

```{r}
m33 <- matrix(c(1,0,0,0,1,1,0,1,1), 3)
plot_binary_matrix(m33)

check_pescado <- function(m33){
  
  resultado <- FALSE  

  # turn around the center, counter-clockwise (CCW), starting at [1,1]
  loop_3x3 <- c(1, 2, 3, 6, 9:7, 4)
  
  # Loop should match exactly 
  pattern <- c(1,0,0,1,1,1,0,0)
  loop_cat <- c(m33[loop_3x3], m33[loop_3x3])
  # loop_sum_cat
  
  # Find the pattern
  idxs.length <- length(loop_cat) - length(pattern) + 1
  for(i in 1:idxs.length){
    
    loop_cat.segment <- loop_cat[i:(i-1+length(pattern))]
    
    if(all(loop_cat.segment == pattern)){
      resultado <- TRUE
      break
    }
  }
  
  return(resultado)
}

check_pescado(m33)
```

### Regla del gancho

Keep

```{r}
m33 <- matrix(c(1,0,1,1,1,1,1,0,0), 3)
plot_binary_matrix(m33)

check_gancho <- function(m33){
  
  m33[2,2] <- 0
  
  sum_condition <- sum(m33) == 5
  
  col_condition <- colSums(m33)[2] == 0
  
  row_condition <- rowSums(m33)[2] == 0
  
  resultado <- sum_condition & (col_condition | row_condition)
  
  return(resultado)
}

check_gancho(m33)
```

### Reglas faltantes / raras

#### Corner 1

```{r}
m33 <- matrix(c(0,0,0,1,1,0,0,1,0), 3)
plot_binary_matrix(m33)
```

#### Corner 2

```{r}
m33 <- matrix(c(1,1,1,0,1,1,0,0,1), 3)
plot_binary_matrix(m33)
```

#### Cruces

```{r}
m33 <- matrix(c(1,0,1,0,1,0,1,0,1), 3)
plot_binary_matrix(m33)

m33 <- matrix(c(0,1,0,1,1,1,0,1,0), 3)
plot_binary_matrix(m33)
```

#### Double fish (bowtie)

```{r}
m33 <- matrix(c(1,1,0,1,1,1,0,1,1), 3)
plot_binary_matrix(m33)
```

#### Regla diagonal (esta mal?)

Kill

```{r}
m33 <- matrix(c(1,0,0,1,1,0,1,1,1), 3)
plot_binary_matrix(m33)

check_diagonal <- function(m33){
  
  resultado <- FALSE  

  # turn around the center, counter-clockwise (CCW), starting at [1,1]
  loop_3x3 <- c(1, 2, 3, 6, 9:7, 4)
  
  # Loop should match exactly 
  pattern <- c(1,0,0,0,1,1,1,1)
  loop_cat <- c(m33[loop_3x3], m33[loop_3x3])
  # loop_sum_cat
  
  # Find the pattern
  idxs.length <- length(loop_cat) - length(pattern) + 1
  for(i in 1:idxs.length){
    
    loop_cat.segment <- loop_cat[i:(i-1+length(pattern))]
    
    if(all(loop_cat.segment == pattern)){
      resultado <- TRUE
      break
    }
  }
  
  return(resultado)
}

check_diagonal(m33)
```

### Rejunte de reglas

```{r}
rules_3x3 <- function(m33, kill_val=0.5, dead_val=0, alive_val=1, no_rule_val = -1){
  
  stopifnot(dim(m33) == c(3,3))
  
  # Border kernels
  if(any(is.na(m33))) return(m33[2,2])
  
  # Check if dead
  if((m33[2,2] == 0)) return(dead_val)
  
  # Check if "straight" segment tip (subtract live first)
  if(((sum(m33) - 1) == 1)) return(kill_val)
  
  # Check if "bent" segment tip
  if(check_bent_tip(m33)) return(kill_val)

  # Check if "protuberant" segment tip
  if(check_protuberance(m33)) return(kill_val)
  
  # Check square
  if(check_square(m33)) return(kill_val)
  
  # Check if surrounded
  if(((sum(m33) - 1) >= 7)) return(alive_val)
  
  # Check if segment
  if(check_segment(m33)) return(alive_val)
  
  # Check if L
  if(check_ele(m33)) return(alive_val)
  
  # Check if S
  if(check_ese(m33)) return(alive_val)
  
  # check if fish
  if(check_pescado(m33)) return(alive_val)
  
  # check if gancho
  if(check_gancho(m33)) return(alive_val)
  
  # Error if no match
  if(no_rule_val == "center") return(m33[2,2])
  return(no_rule_val)
}
```

## Iterador de autómata falopa

Get square by center coordinate:

```{r}
get_3x3 <- function(m,i.row=2,j.col=2){
  
  if(i.row < 2) stop("Indexes for rows and columns must be 2 or greater.")
  if(j.col < 2) stop("Indexes for rows and columns must be 2 or greater.")
  
  m[(i.row-1):(i.row+1), (j.col-1):(j.col+1)]
}
```


```{r}
conway_erode <- function(mask, kill_val = 0.5, no_rule_val = -1){
  mask.result <- mask
  mask.paded <- mat_pad(mask)
  
  for(i.row in 1:dim(mask)[1]){
    for(j.col in 1:dim(mask)[2]){
      
      mat_3x3 <- get_3x3(m = mask.paded, 
                         i.row = i.row+1, 
                         j.col = j.col+1)
      
      mask.result[i.row, j.col] <- rules_3x3(mat_3x3, kill_val = kill_val, no_rule_val = no_rule_val)
        
    } 
  }
  
  return(mask.result)
}
```

## Other 8x8 tests

### Test 2: more patterns

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```


```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_001.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_006.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_002.tif", binarize = T)
mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_003.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mask)
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_004.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mask)
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_005.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mask)
plot_binary_matrix(mask.result)
```

### Test 3: iteration

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix.tif", binarize = T)

iterate_erode <- function(mask, ..., .dir=tempdir()){
  
  if(file.exists(.dir)) unlink(.dir, recursive = T)
  dir.create(.dir, showWarnings = F, recursive = T)
  
  png(filename = tempfile(pattern = paste(0, "-mask"), tmpdir = .dir, fileext = ".png"))
  print(plot_binary_matrix(mask))
  dev.off()

  for(i in 1:10){
    mask.post <- conway_erode(mask, kill_val = 0, no_rule_val = "center", ...)
    
    if(identical(mask, mask.post)){
      break
    } else {
      png(filename = tempfile(pattern = paste(i, "-mask"), tmpdir = .dir, fileext = ".png"))
      print(plot_binary_matrix(mask.post))
      dev.off()
      
      mask <- mask.post
    }
  }
}

iterate_erode(mask, .dir = "/tmp/test1")

knitr::include_graphics(dir("/tmp/test1", full.names = T))
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/48pix.tif", binarize = T)

iterate_erode <- function(mask, ..., .dir=tempdir()){
  
  if(file.exists(.dir)) unlink(.dir, recursive = T)
  dir.create(.dir, showWarnings = F, recursive = T)
  
  # png(filename = tempfile(pattern = paste(0, "-mask"), tmpdir = .dir, fileext = ".png"))
  print(plot_binary_matrix(mask))
  # dev.off()

  for(i in 1:10){
    mask.post <- conway_erode(mask, kill_val = 0, no_rule_val = "center", ...)
    
    if(identical(mask, mask.post)){
      break
    } else {
      # png(filename = tempfile(pattern = paste(i, "-mask"), tmpdir = .dir, fileext = ".png"))
      print(plot_binary_matrix(mask.post))
      # dev.off()
      
      mask <- mask.post
    }
  }
}

iterate_erode(mask, .dir = "/tmp/test2")

knitr::include_graphics(dir("/tmp/test2", full.names = T))
```


```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_001.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_002.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_003.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_004.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_005.tif", binarize = T)
iterate_erode(mask)
```

## Proxima idea

* Buscar el patron 101 / 1001 / 10001 / etc. en la secuencia de ir alrededor al pixel central, para encontrar si el pixel del medio conecta dos pixeles de alrededor.
* Si se muere, pasar la vida al de al lado (ojo con oscilaciones, debe haber un criterio de terminacion, quizas decir que con un poquito de vida queda totalmente vivo para la comparación con la iteracion anterior).
