---
title: "Edge erosion for Cell-ID masks"
---

```{r setup}
knitr::opts_chunk$set(message = F)
knitr::opts_knit$set(root.dir = here::here())
# knitr::opts_chunk$set(cache = T)

library(utiles)
library(dplyr)
library(ggplot2)
# library(tidyr)
# library(purrr)
# library(plotly)
# library(renv)

library(mmand)
library(magick)
library(igraph)
```


## Load data

```{r}
pics <- dir(path = "testings/erosion_dilation/test_imgs/", full.names = T, pattern = "tif")

pic_mats <- lapply(pics, read_tif_to_matrix)

# lapply(pic_mats, plot_matrix)
```

Me gusta el smooth con `sigma = c(0.5,0.5)`:

```{r}
m <- pic_mats[[2]]
m.smooth <- gaussianSmooth(x = m, sigma = c(0.5,0.5))
plot_matrix(m)
plot_matrix(m.smooth)
```


```{r}
k <- shapeKernel(width = rep(2,2), type="diamond")  #type = c("box", "disc", "diamond")

plot_matrix(m) + ggtitle("Original")
plot_matrix(erode(m,k)) + ggtitle("Erode")
plot_matrix(dilate(m,k)) + ggtitle("Dilate")

plot_matrix(m.smooth) + ggtitle("Smooth")
plot_matrix(erode(m.smooth,k)) + ggtitle("Smooth Erode")
plot_matrix(dilate(m.smooth,k)) + ggtitle("Smooth Dilate")

plot_matrix(dilate(m,k) - erode(m,k)) + ggtitle("DE over original matrix")
plot_matrix(dilate(m.smooth,k) - erode(m.smooth,k)) + ggtitle("DE over smooth matrix")
```

## Automata falopa v2

El problema de este es que en muchas geometrias de frontera,
el pixel del medio es un camino mas corto entre otros dos localmente
pero que esta en la frontera que queremos borrar!

Entonces borra corners bien, pero fronteras mal :(

### Regla vecinos conectados

Kill

```{r}
m33 <- matrix(c(0,0,1,1,1,0,1,1,0), 3)
m33 <- matrix(c(0,0,0,1,1,0,1,1,0), 3)
plot_binary_matrix(m33)
```


```{r}
check_neighs <- function(m33, max_dist=sqrt(2)){
  m33[2,2] <- 0
  # plot_binary_matrix(m33)
  
  # If there are less than 2 neighbours, this one may be killed.
  if(sum(m33) < 2) return(TRUE)
  
  dists <- 
    matrix_to_df(m33) %>% filter(pix_value == 1) %>% select(-pix_value) %>% 
    dist() %>% as.matrix()
  dists
  
  diag(dists) <- NA
  # dists[upper.tri(dists)] <- NA
  dists
  
  closest_neigh_dist <- dists %>% apply(MARGIN = 2, function(x) min(x, na.rm=T))
  
  # If TRUE we *can* kill this pixel,
  # because neighbours do not depend on this one to be connected directly
  # i.e. they are all adyacent to at least one of other non-central pixels
  closest_neigh_dist.condition <- all(closest_neigh_dist <= max_dist)
  
  # ¿But should we kill it? A 1-filled matrix will say "you may kill"
  # Even though we should not kill it, obviously
  # To decide, I can use the erode function or check for the center's betweenness
  return(closest_neigh_dist.condition)
}

plot_binary_matrix(m33) + ggtitle("Delete?", check_neighs(m33))
```

Tambien queria chequear si el pixel del medio esta en un shortest path o no.

Si no lo estaba, no ayuda a conectar pixeles realmente, y se deberia borrar.

```{r}
m33.yes <- matrix(c(1,0,0,1,1,0,1,0,0), 3)
m33.no <- matrix(c(0,1,0,0,1,1,0,1,0), 3)
# plot_binary_matrix(m33)

weighted_shortest_paths.condition <- function(allpaths, v){
  sapply(allpaths, function(p) v %in% p) %>% any()
}

has.weighted_shortest_paths <- function(cgraph, 
                                        v=which(V(cgraph)$coords == "2,2"),
                                        condition.function=weighted_shortest_paths.condition){
  #initialize a list to populate with all the shortest paths in the graphy
  # allpaths <- c()
  allpaths <- list()
  
  n.verts <- 1:length(V(cgraph))
  
  #Assuming this is an undirected graph, we don't want to calculate both a %--% b and b %--% a  
  for(i in n.verts){
    for(j in n.verts){
      if(i != v & j != v & i < j){
        shortest_path_options <- all_shortest_paths(cgraph, i, j)$res
        #sometimes there are multiple shortest paths, we will include them all
        # allpaths <- append(allpaths, sum(sapply(shortest_path_options, length)))
        allpaths <- append(allpaths, shortest_path_options)
      }
    }
  }
  
  condition <- condition.function(allpaths, v)
  
  return(condition)
}



check_path <- function(m33, max_dist=sqrt(2)){
  # plot_binary_matrix(m33)
  
  xy <- matrix_to_df(m33) %>% 
    filter(pix_value > 0) %>% 
    select(x,y)
  
  dmat <- as.matrix(
    dist(xy, upper = T, diag = T)
  )
  
  # cmat <- dmat <= max_dist
  
  dmat[dmat > max_dist] <- 0
  cmat <- dmat
  
  cgraph <- igraph::graph.adjacency(cmat, mode = "undirected", diag = F, 
                                    weighted = "weight",
                                    add.colnames = "id"
                                    )
  V(cgraph)$coords <- apply(xy,1,paste,collapse=",")
  V(cgraph)$color <- ifelse(V(cgraph)$coords == "2,2", "lightblue", "orange")
  # plot.igraph(cgraph, edge.color=E(cgraph)$weight*10, edge.width=E(cgraph)$weight*10)
  
  # TRUE ~ Can kill
  # FALSE ~ Do not kill
  condition <- !has.weighted_shortest_paths(cgraph)
  # center_betweenness <- betweenness(cgraph, V(cgraph)[V(cgraph)$coords == "2,2"])
  # center_betweenness.condition <- center_betweenness == 0
  
  return(condition)
}

plot_binary_matrix(m33.yes) + ggtitle("Delete m33.yes?", check_path(m33.yes))
plot_binary_matrix(m33.no) + ggtitle("Delete m33.no?", check_path(m33.no))
```

```{r}
# c.nodes <- vertex_attr(cgraph) %>% as.data.frame()
# c.edges <- as.data.frame(igraph::ends(cgraph, es = E(cgraph)))
# c.edges[,3] <- edge_attr(cgraph)$weight
# names(c.edges) <- c("from", "to", "weight")
# 
# c.nodes
# c.edges
```


```{r}
# abs_betweenness <- function(x){
#   temp_bool <- sapply(allpaths, function(y){
#     (x %in% y) & (head(y, 1) != x) & (tail(y, 1) != x)
#   })
#   length(allpaths[temp_bool])
# }
# 
# abs_betweenness()
```


### Integración de reglas

```{r}
rules_3x3 <- function(m33, kill_val=0.5, dead_val=0, alive_val=1, no_rule_val = -1,
                      default_val = m33[2,2]){
  
  plot_binary_matrix(m33)
  stopifnot(dim(m33) == c(3,3))
  
  if(m33[2,2] == dead_val) return(dead_val)
  
  if(any(is.na(m33))) return(default_val)
  
  kill.conditon <- check_path(m33) & check_neighs(m33)
  
  result <- ifelse(kill.conditon, kill_val, alive_val)
  
  return(result)
}
```

## Iterador de autómata falopa

Get square by center coordinate:

```{r}
get_3x3 <- function(m,i.row=2,j.col=2){
  
  if(i.row < 2) stop("Indexes for rows and columns must be 2 or greater.")
  if(j.col < 2) stop("Indexes for rows and columns must be 2 or greater.")
  
  m[(i.row-1):(i.row+1), (j.col-1):(j.col+1)]
}
```


```{r}
conway_erode <- function(mask, kill_val = 0.5, no_rule_val = -1){
  mask.result <- mask
  mask.paded <- mat_pad(mask)
  
  # i.row <- 2
  # j.col <- 6
  
  for(i.row in 1:dim(mask)[1]){
    for(j.col in 1:dim(mask)[2]){
      
      # print(paste(i.row, j.col))
      
      mat_3x3 <- get_3x3(m = mask.paded, 
                         i.row = i.row+1, 
                         j.col = j.col+1)
      
      mask.result[i.row, j.col] <- rules_3x3(m33 = mat_3x3, 
                                             kill_val = kill_val)
        
    } 
  }
  
  return(mask.result)
}
```

## Other 8x8 tests

### Test 2: more patterns

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix.tif", binarize = T)
plot_binary_matrix(mask)

mask.result <- conway_erode(mask)

# plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```


```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_001.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_006.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_002.tif", binarize = T)
mask.result <- conway_erode(mask)

plot_binary_matrix(mat_pad(mask))
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_003.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mask)
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_004.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mask)
plot_binary_matrix(mask.result)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_005.tif", binarize = T)

mask.result <- conway_erode(mask)

plot_binary_matrix(mask)
plot_binary_matrix(mask.result)
```

### Test 3: iteration

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix.tif", binarize = T)

iterate_erode <- function(mask, ..., .dir=tempdir()){
  
  if(file.exists(.dir)) unlink(.dir, recursive = T)
  dir.create(.dir, showWarnings = F, recursive = T)
  
  png(filename = tempfile(pattern = paste(0, "-mask"), tmpdir = .dir, fileext = ".png"))
  print(plot_binary_matrix(mask))
  dev.off()

  for(i in 1:10){
    mask.post <- conway_erode(mask, kill_val = 0, no_rule_val = "center", ...)
    
    if(identical(mask, mask.post)){
      break
    } else {
      png(filename = tempfile(pattern = paste(i, "-mask"), tmpdir = .dir, fileext = ".png"))
      print(plot_binary_matrix(mask.post))
      dev.off()
      
      mask <- mask.post
    }
  }
}

iterate_erode(mask, .dir = "/tmp/test1")

knitr::include_graphics(dir("/tmp/test1", full.names = T))
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/48pix.tif", binarize = T)
plot_binary_matrix(mask)

iterate_erode <- function(mask, ..., .dir=tempdir()){
  
  if(file.exists(.dir)) unlink(.dir, recursive = T)
  dir.create(.dir, showWarnings = F, recursive = T)
  
  # png(filename = tempfile(pattern = paste(0, "-mask"), tmpdir = .dir, fileext = ".png"))
  print(plot_binary_matrix(mask))
  # dev.off()

  for(i in 1:10){
    mask.post <- conway_erode(mask, kill_val = 0, no_rule_val = "center", ...)
    
    if(identical(mask, mask.post)){
      break
    } else {
      # png(filename = tempfile(pattern = paste(i, "-mask"), tmpdir = .dir, fileext = ".png"))
      print(plot_binary_matrix(mask.post))
      # dev.off()
      
      mask <- mask.post
    }
  }
}

iterate_erode(mask, .dir = "/tmp/test2")

knitr::include_graphics(dir("/tmp/test2", full.names = T))
```


```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_001.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_002.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_003.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_004.tif", binarize = T)
iterate_erode(mask)
```

```{r}
mask <- utiles::read_tif_to_matrix("testings/edge_kernels/8pix_005.tif", binarize = T)
iterate_erode(mask)
```

## Proxima idea

Quizas integrando `contract_by_cointainment` de `erosion_dilation.Rmd` logre algo.

Ese si hacia bien el erosion sobre las fronteras (rectas al menos).
Hay que arreglarlo para que coma diagonales tmb (todo sin comer los "parents").

Una facil seria erosionar "a mano" pixeles de diagonales (siempre que
no sean parents de otro) por patrón hardcoreado.
